---
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(forcats))
suppressPackageStartupMessages(library(scales))
```

## Context

I scraped all the historical data from the Puerto Rico's Electoral Commission webpage to make it accesible in a way that can be combined with Census data, as well as other socio-economic indicators to better explain political patterns in Puerto Rico. In the next month I will add as well the elections of 2004, 2008, 2012, 2016, but with a more granual level than municipalities. I parsed the names of the parties I knew, but the other ones stayed with the code assigned by the CEE, if you know them, and can provide a reference please send me an email at: iflores.siaca@gmail.com

> The GitHub repository with all the code for this project is available here: [https://github.com/ian-flores/PR_Elections_Historical_Data](https://github.com/ian-flores/PR_Elections_Historical_Data)

> Disclaimer: The data contained in this page was pulled from the Puerto Rico's Electoral Commission webpage. Where there to be any discrepancies, please refer to their page.

### Data Available

- Governor *(1948 - 2000)*
- Resident Commissioner *(1932 - 1944) & (1980 - 2000)*
- Mayors *(1976 - 2000)*

### Governor Data Available

> This data is available to download here: [Governor_CSV](https://github.com/ian-flores/PR_Elections_Historical_Data/blob/master/data/elecciones_generales/gobernador/gobernador.csv)

```{r, include = TRUE, fig.width=10, fig.height=6}
data <- read_csv('data/elecciones_generales/gobernador/gobernador.csv')

ggthemr::ggthemr('fresh')

data %>%
    group_by(year_data, partido) %>%
    summarise(votos = sum(votos, na.rm = T)) %>%
    filter(votos > 75000) %>%
    mutate(partido = fct_drop(partido)) %>%
    rename(year = year_data, votes = votos, party = partido) %>%
    ggplot(aes(x = year, y = votes, colour = fct_reorder2(party, year, votes))) +
        geom_line(aes(group = party)) +
    labs(x = 'Year', y = 'Votes', 
         colour = 'Political Party',
         title = 'Governor Election Results',
         subtitle = 'Data from 1948-2000',
         caption = 'Only parties with more than 75,000 votes each year are displayed') +
    scale_y_continuous(labels = comma_format())
```

```{r, eval=FALSE}
knitr::kable(head(data, n = 10), col.names = c('Municipality', 'Year', 'Political Party', 'votes'))
```

### Resident Commissioner Data Available

> This data is available to download here: [Commissioner_CSV](https://github.com/ian-flores/PR_Elections_Historical_Data/blob/master/data/elecciones_generales/comisionado_residente/comisionado_residente.csv)

```{r, include = TRUE, fig.width=10, fig.height=6}
data <- read_csv('data/elecciones_generales/comisionado_residente/comisionado_residente.csv')

ggthemr::ggthemr('fresh')

data %>%
    group_by(year_data, partido) %>%
    summarise(votos = sum(votos, na.rm = T)) %>%
    filter(year_data > 1979) %>%
    filter(votos > 75000) %>%
    mutate(partido = fct_drop(partido)) %>%
    ggplot(aes(x = year_data, y = votos, colour = fct_reorder2(partido, year_data, votos))) +
        geom_line(aes(group = partido)) +
    labs(x = 'Year', y = 'Votes', 
         colour = 'Political Party',
         title = 'Resident Commisioner Results',
         subtitle = 'Data from 1980-2000',
         caption = 'Only parties with more than 75,000 votes each year are displayed') +
    scale_y_continuous(labels = comma_format())
```

```{r, eval=FALSE}
knitr::kable(head(data, n = 10), col.names = c('Municipality', 'Year', 'Political Party', 'votes'))
```

### Mayors Data Available

> This data is available to download here: [Mayor_CSV](https://github.com/ian-flores/PR_Elections_Historical_Data/blob/master/data/elecciones_generales/alcalde/alcalde.csv)

```{r, include = TRUE, fig.width=10, fig.height=6}
data <- read_csv('data/elecciones_generales/alcalde/alcalde.csv')


data %>%
    group_by(year_data, partido) %>%
    summarise(votos = sum(votos, na.rm = T)) %>%
    filter(votos > 25000) %>%
    mutate(partido = fct_drop(partido)) %>%
    rename(year = year_data, votes = votos, party = partido) %>%
    ggplot(aes(x = year, y = votes, colour = fct_reorder2(party, year, votes))) +
        geom_line(aes(group = party)) +
    labs(x = 'Year', y = 'Votes', 
         colour = 'Political Party',
         title = 'Mayor Results',
         subtitle = 'Data from 1976-2000',
         caption = 'Only parties with more than 25,000 votes each year are displayed') +
    scale_y_continuous(labels = comma_format())
```

```{r, eval=FALSE}
knitr::kable(head(data, n = 10), col.names = c('Municipality', 'Year', 'Political Party', 'votes'))
```